---
title: "project1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(CBPS)
library(clusterSEs)
library(DescTools)
library(data.table)
library(dplyr)
library(geosphere)
library(ggplot2)
library(glue)
library(haven)
library(plm)
library(readxl)
library(RColorBrewer)
library(stargazer)
library(tidyverse)
library(xtable)
library(kknn)
library(rlist)
```

# data processing

## housing price indices
Federal Housing Finance Agency:
FHFA house price index, all-transactions indexes ( estimated using sales prices and appraisal data)
1975-2021, quarterly estimates
https://www.fhfa.gov/DataTools/Downloads/Documents/HPI/HPI_AT_state.csv

annual estimates
https://www.fhfa.gov/DataTools/Downloads/Documents/HPI/HPI_AT_BDL_state.xlsx
For tracking and feedback purposes, please cite: Bogin, Alexander N., Doerner, William M. and Larson, William D. (2019). "Local House Price Dynamics: New Indices and Stylized Facts". Real Estate Economics, volume 47, issue 2, pages 365-398. 

problem: the FHFA index is only for comparison within states.each satets starts at the starting year as 100.
for comparison cross state, the growth rate can be used.makes sense to use year on year change rate instead of quarterly rates, less noisy.

MHS data:
Average sales price by state. only recorded newly manufactured homes sold or apprraised.2014-2020

Unemployment data from BLS
https://www.bls.gov/web/laus/ststdsadata.zip

## variable: HPI growth rate per state per year, 2000-2019, HPI_yr,  HPI_yr_l
```{r}
HPI_yr <- read.csv("HPI_state_year.csv")
colnames(HPI_yr)<- c("state","abbr","FIPS","year","rate","HPI","HPI90","HPI00")

HPI_yr <- HPI_yr[which(HPI_yr$year>1999 & HPI_yr$year<2020),]
HPI_yr <- HPI_yr[,c(2,4,5)] # change rates in col5, levels in col6, HPI level starts at 100 for all states on 1975 levels

HPI_yr_l <- HPI_yr # keep the long form for panel model
HPI_yr_l <- HPI_yr_l %>% mutate(st_yr=paste0(abbr,year))

HPI_yr <- HPI_yr%>% pivot_wider(.,names_from = "abbr",values_from = "rate")
```



## create panel weght matrix, policy diffusion net pol_netw

```{r}
# policy diffusion network 1960-2009 Desmarais 2015, no panel for 1995-2015 use the year 1995? 2009?, used 6 year lagged panel 1989-2009?
# note pol_net state abbreviation are alphabetically sorted by the abbr not the full statenames!!!!! need to be matched with usual order before carry on

# on 01.14 the csv has been modified so that the first column is the destination state, the second column the origin state. maks later multiplication easier (suming over rows to get a col vec)

pol_net <- read.csv("dhb2015apsr-networks.csv")
pol_net <- pol_net[,1:4] # use the 400 edged in 10 years network

statename_order <- read_xlsx("statename.xlsx")
temp <- do.call("cbind",replicate(51,statename_order,simplify=FALSE)) 
colnames(temp) <- c("state","AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY")
temp <- temp %>% pivot_longer(.,!state,names_to = "state_02",values_to = "fooo" )
temp <- temp %>% mutate(OD=paste0(state,state_02))
temp <- temp[,-c(3)]
order<- temp# save the order for later, the order of 2500 OD pairs

# function to get the pol_net for a specific year
get_pol_net <- function(yr) {
  # input th year, output pol_net for the year in wide format
  net <- pol_net %>% filter(.,year==yr)
  net <-net[,c(1,2,4)]
  net[is.na(net)] <- 0

  # fixing the ordering of states
  net <- net %>% mutate(OD=paste0(state_01,state_02))
  net <- left_join(order,net[,c(3,4)],by=c("OD")) # order fixed

  net <- net[,-c(3)]
  net <- net %>% pivot_wider(.,names_from = "state_02",values_from = "src_10_400")
}

# not using list for convenience
pol_net0 <- get_pol_net(2000)
pol_net1 <- get_pol_net(2001)
pol_net2 <- get_pol_net(2002)
pol_net3 <- get_pol_net(2003)
pol_net4 <- get_pol_net(2004)
pol_net5 <- get_pol_net(2005)
pol_net6 <- get_pol_net(2006)
pol_net7 <- get_pol_net(2007)
pol_net8 <- get_pol_net(2008)
pol_net9 <- get_pol_net(2009)

col1 <- pol_net0[,1]
```

## information diffusion net (geo proximity) 
```{r}
# state coordinates
coord <- read.csv("statelatlong.csv")
coord <- coord[-9,]
statename <- c("AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY")

# distance matrix from coordinates
dist.m <- distm(coord[2:1],coord[2:1],fun=distVincentySphere) # long first, lat second, great circle distance assuming earth is spherical
colnames(dist.m) <- statename

# defining congruent neighbors
# for each state find 6 nearest neighbours, 
# in each column of nbr the 6 nearest state to the state in the col name
nbr <- NULL
for (i in 1:50) {
  t <- as.vector(dist.m[,i])
  t <- sort(t,decreasing=TRUE)
  tt <- t[44] # choose 6 nearest neighbour
  ttt <- as.data.frame(dist.m[,i])
  colnames(ttt)<- "val"
  ttt<- ifelse(ttt$val>tt,0,1 ) 

  nbr <- cbind(nbr,ttt)
}
colnames(nbr) <- statename

# in transposed nbr, each row as the nearest six states to the state in the row name
nbr <- t(nbr)
diag(nbr) <- 0
```

## these 50*50 nbrnpol weights are to be used to form HPI_nbr_lag
```{r}
# among nearest neighbours, keep those that are not in the pol_net neighbours of the year
# inverted pol neighbour net, policy neighbours has 0, non neighbours has 1

get_nbrnpol <- function(polnet) {
  # input 50*50 wide format pol_net for a gven yeat 
  # output a 50*50 nbr weight matrix for a given year without these pol neighbours
  
  net_inv <- polnet[,-1]
  mat <- matrix(1,50,50)
  net_inv <- mat+net_inv
  net_inv <- ifelse(net_inv==2,0,1)

  nbr_npol <- nbr*net_inv # multiply so that geo neighbour elements also in pol net will zero out
  
  # standardise rows
  div<- rowSums(nbr_npol)
  div <- do.call("cbind",replicate(50,div,simplify=FALSE))
  nbr_npol <- nbr_npol/div
  
}

nbr_npol0<-get_nbrnpol(pol_net0)
nbr_npol1<-get_nbrnpol(pol_net1)
nbr_npol2<-get_nbrnpol(pol_net2)
nbr_npol3<-get_nbrnpol(pol_net3)
nbr_npol4<-get_nbrnpol(pol_net4)
nbr_npol5<-get_nbrnpol(pol_net5)
nbr_npol6<-get_nbrnpol(pol_net6)
nbr_npol7<-get_nbrnpol(pol_net7)
nbr_npol8<-get_nbrnpol(pol_net8)
nbr_npol9<-get_nbrnpol(pol_net9)

# make them a list
nbr_netss <- list(nbr_npol0,nbr_npol1,nbr_npol2,nbr_npol3,nbr_npol4,nbr_npol5,nbr_npol6,nbr_npol7,nbr_npol8,nbr_npol9,nbr_npol9,nbr_npol9,nbr_npol9,nbr_npol9,nbr_npol9,nbr_npol9,nbr_npol9,nbr_npol9,nbr_npol9,nbr_npol9)
```


## exclude geo nbr from pol net nbr, these 50*50 polnet are to be used to form HPI_pol_lag
```{r}

get_pol_net2 <- function(polnet) {
  # input 50*50 polnet 
  # output the polnet exclusing nearest 6 geo nbrs

  mat <- matrix(1,50,50)
  nbr_inv <- mat+nbr
  nbr_inv <- ifelse(nbr_inv==2,0,1)
  
  col1 <- polnet[,1]
  polnet <- polnet[,-c(1)]
  polnet <- polnet*nbr_inv

  # row standardise pol_net_2009w for multiplication use
  div<- rowSums(polnet)
  div <- do.call("cbind",replicate(50,div,simplify=FALSE))
  polnet <- polnet/div
}

pol_net0 <- get_pol_net2(pol_net0)
pol_net1 <- get_pol_net2(pol_net1)
pol_net2 <- get_pol_net2(pol_net2)
pol_net3 <- get_pol_net2(pol_net3)
pol_net4 <- get_pol_net2(pol_net4)
pol_net5 <- get_pol_net2(pol_net5)
pol_net6 <- get_pol_net2(pol_net6)
pol_net7 <- get_pol_net2(pol_net7)
pol_net8 <- get_pol_net2(pol_net8)
pol_net9 <- get_pol_net2(pol_net9)

# make them a list
pol_netss <- list(pol_net0,pol_net1,pol_net2,pol_net3,pol_net4,pol_net5,pol_net6,pol_net7,pol_net8,pol_net9,pol_net9,pol_net9,pol_net9,pol_net9,pol_net9,pol_net9,pol_net9,pol_net9,pol_net9,pol_net9)
```


## variable: lagged HPI growth in policy net sources HPI_pol_lag_l
```{r}
# make 50*20 matrix for each year HPI rate, state names in colnames, rows as repetitions.
# pol_net_2009w2 is row standardised, pol_net_2009w is not. choose either one
HPI_yr2 <- HPI_yr[,-c(1,10)] # delete year index and DC

############################
# function:  get_spatial_lag()
############################

get_spatial_lag<- function(mat,n,w){
  # input: mat: thet data matrix 20*50 20 rows for the 20 years, 50 cols for the 50 states
  #        ra:  1*50, selected by n from mat, data (HPI) by state by single year, the n row of HPI_yr2 , 50 states in colnames, 1 rows for 1 years
  #        w:   50*50  state to state adjacency matrix (or spatial weight matrix)
  # output: 50*1 matrix, 50 rows for 50 states, 1 cols for 1 years.
  ra <- mat[n,]
  temp <- do.call("rbind",replicate(50,ra,simplify=FALSE)) # repeat 50 rows
  temp <- temp*w # multiply HPI by weights
  b <- as.data.frame(rowSums(temp))
}

bb<- NULL
for (i in 1:20) {
 bb[i] <-get_spatial_lag(HPI_yr2,i,pol_netss[[i]])
}

bb <- list.cbind(bb)

colnames(bb) <- c("2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018","2019")
bb <- cbind(col1,bb) 

HPI_pol_lag_l <- bb %>% pivot_longer(.,!state,names_to="year",values_to="HPI_pol_lag") # shape to long format for panel model
HPI_pol_lag_l <- HPI_pol_lag_l %>% mutate(.,st_yr=paste0(state,year))
```

## variable: lagged HPI growth in geo neighbours
```{r}
# similar procedure with pol_lag, using spatial contiguity matrix instead
bb<- NULL
for (i in 1:20) {
 bb[i] <-get_spatial_lag(HPI_yr2,i,nbr_netss[[i]])
}

bb <- list.cbind(bb)

colnames(bb) <- c("2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018","2019")
bb <- cbind(col1,bb) 

HPI_nbrnpol_l <- bb %>% pivot_longer(.,!state,names_to="year",values_to="HPI_nbrnpol_lag") # shape to long format for panel model
HPI_nbrnpol_l <- HPI_nbrnpol_l %>% mutate(.,st_yr=paste0(state,year))
```

## variable: WX unemployment rate as IV
```{r}
# shape empdata to wide format 20*50
empdata <- read_csv("empdata.csv")
empdata <- empdata[,c(2,3,5)]
empdata <- empdata %>% pivot_wider(.,names_from="state",values_from="emp")
empdata <- empdata[,-1] # lose the year col, now same shape as HPI_yr2

# pol net emp
bb<- NULL
for (i in 1:20) {
 bb[i] <-get_spatial_lag(empdata,i,pol_netss[[i]])
}

bb <- list.cbind(bb)

colnames(bb) <- c("2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018","2019")
bb <- cbind(col1,bb) 

emp_pol_lag_l <- bb %>% pivot_longer(.,!state,names_to="year",values_to="emp_pol_lag") # shape to long format for panel model
emp_pol_lag_l <- emp_pol_lag_l %>% mutate(.,st_yr=paste0(state,year))

# geo nbr emp
bb<- NULL
for (i in 1:20) {
 bb[i] <-get_spatial_lag(empdata,i,nbr_netss[[i]])
}

bb <- list.cbind(bb)

colnames(bb) <- c("2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018","2019")
bb <- cbind(col1,bb) 

emp_nbr_l <- bb %>% pivot_longer(.,!state,names_to="year",values_to="emp_nbr_lag") # shape to long format for panel model
emp_nbr_l <- emp_nbr_l %>% mutate(.,st_yr=paste0(state,year))

```

## bank/tax policy counts
```{r}
# state legislation by policy category, 1991-2018
# choose the number of bank (banking, real estimate) policies in state/year, as proxy of policy influence

state_legislation<-read_dta("Garlick_StateLegislationPolicy.0.1.dta")

#state <- unique(state_legislation$state) # 50 states, not DC
#policy_cat <- unique(state_legislation$polname)
# 20 cat "Bank"  "Construction" "Tax Policy" "Communication" "Manufacturing" "Law" "Agriculture" "Transportation" "Environment" "Welfare"  "Health" "Sports and Recreation" "Religion"  "Small Business"   "Insurance"  "Education" "Local Government" "Police and Fire" "Civil Rights" "Agriculture"
#write.csv(state_legislation,file="statelegislation.csv")

############################
# function: get_policy_count
############################

# input: string, of the policy name in state_legislation
# output: 1000*4  long format state by year by count of the chosen policy

get_policy_count <- function (nam) {
  temp <- state_legislation[,c(1,2,7)] %>% filter(.,polname==as.character(nam))%>% mutate(.,state_year=paste0(state,year1))
  count1 <- table(temp$state_year)
  temp <- unique(temp)
  temp <- cbind(temp[,c(1,3)],count1)
  temp <- temp %>% rename(.,"policy_count"="Freq")
}

# bank (banking and real estate related)
policy_bank <- get_policy_count("Bank")

# other policies
policy_tax <- get_policy_count("Tax Policy")
policy_const <-get_policy_count("Construction")

# take to log form
policy_bank <- policy_bank %>% mutate_at(.,"policy_count",log)
policy_tax <- policy_tax %>% mutate_at(.,"policy_count",log)
policy_const <- policy_const %>% mutate_at(.,"policy_count",log)
#rm(state_legislation)
```

## variable: neighbour tax policy counts in policy nets, to instrument for policy net housing prices
```{r}
############################
# function: get_policy_lag
############################

# this only works for the policy counts data! It backdates the counts.

# input:   W    1000*4 state_year policy count matrix, the policy count in each given state_year
#          net  list[20] list of 20 50*50 weight matrices, pol_netss or nbr_netss
# output:  1000*4 for each state_year, the number of policies passed in its policy net upstream states.
# uses the NOT row standardised policy net adjacency matrix

get_policy_lag <- function(w,net) {
  
  w <- w[,c(1,2,4)]
  w <- w[which(w$year1>1998 & w$year1<2019),] # get one year lag policy counts for year 2000-2019
  w <- w %>% mutate(plus1=1)
  w <- w %>% mutate(year1=year1+plus1) # pad the year with 1, easier when merge data
  w <- w %>% mutate(st_yr=paste0(state,year1))

  # ordering states by usual order
  temp <- do.call("cbind",replicate(21,statename_order,simplify=FALSE)) 
  colnames(temp) <- c("state","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018","2019")
  temp <- temp %>% pivot_longer(.,!state,names_to = "year",values_to = "fooo" )
  temp <- temp %>% mutate(OD=paste0(state,year))
  temp <- temp[,-c(3)]
  #order2<- temp
  w <- left_join(temp,w[,c(3,5)],by=c("OD"="st_yr"))
  w[is.na(w)] <- 0

  w_w <- w[,c(1,2,4)] %>% pivot_wider(.,names_from = "state", values_from = "policy_count")
  w_w <-w_w[,-c(1)]

  # get policy lag in policy net neighbours
  w_lag <- NULL
  for (i in 1:20) {
    temp <- w_w[i,]
    temp <- do.call("rbind",replicate(50,temp,simplify=FALSE)) # repeat 50 rows
    b <- net[[i]]
    temp <- temp*b # multiply HPI by weights, each row is for a certain state the HPI weighted by its policy net neighbours
    b <- as.data.frame(rowSums(temp))
    w_lag[i] <- b
  }

  w_lag <- as.data.frame(w_lag)
  colnames(w_lag) <- c("2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018","2019")

  w_lag <- cbind(col1,w_lag)

  w_lag_l <- w_lag %>% pivot_longer(.,!state,names_to="year",values_to="pol_lag") # shape to long format for panel model
  w_lag_l <- w_lag_l%>%mutate(st_yr=paste0(state,year))
  
}

bank_pol_lag_l<- get_policy_lag(policy_bank,pol_netss)
colnames(bank_pol_lag_l)<- c("state","year","bank_pol_lag","st_yr")

tax_pol_lag_l<- get_policy_lag(policy_tax,pol_netss)
colnames(tax_pol_lag_l)<- c("state","year","tax_pol_lag","st_yr")

const_pol_lag_l<- get_policy_lag(policy_const,pol_netss)
colnames(const_pol_lag_l)<- c("state","year","const_pol_lag","st_yr")
```

## variable: neighbour tax policy counts in geo nbrs, to instrument for geo nbr housing prices
```{r}

bank_nbr_lag_l<- get_policy_lag(policy_bank,nbr_netss)
colnames(bank_nbr_lag_l)<- c("state","year","bank_nbr_lag","st_yr")

tax_nbr_lag_l<- get_policy_lag(policy_tax,nbr_netss)
colnames(tax_nbr_lag_l)<- c("state","year","tax_nbr_lag","st_yr")

const_nbr_lag_l<- get_policy_lag(policy_const,nbr_netss)
colnames(const_nbr_lag_l)<- c("state","year","const_nbr_lag","st_yr")
```



## controls
 GDP 1997-2919, population 2000-2019, 
 housing unit estimate 2000-2010, 2010-2019: https://www.census.gov/data/tables/time-series/demo/popest/intercensal-2000-2010-housing-units.html
 
```{r}
# import files
num_hs1 <- read_xls("st-hs-unit-2000-10.xls")
num_hs2 <- read_xlsx("st-hs-unit-2010-19.xlsx")
num_hs <- left_join(num_hs1,num_hs2)

pop1 <- read.csv("st-pop-2000-10.csv")
pop2 <- read_xlsx("st-pop-2010-19.xlsx")
pop <- left_join(pop1,pop2)
colnames(pop) <- colnames(num_hs)

gdp <- read.csv("GDP_state_1997_2019.csv")
gdp <- gdp[,-c(2:4)]
colnames(gdp) <- colnames(num_hs)

# merge and format control files
num_hs_l <- num_hs %>% pivot_longer(.,!state,names_to = "year",values_to = "num_house")
num_hs_l <- num_hs_l %>% mutate(.,st_yr=paste0(state,year))

pop_l <- pop %>% pivot_longer(.,!state,names_to = "year",values_to = "pop")
pop_l <- pop_l %>% mutate(.,st_yr=paste0(state,year))

gdp_l <- gdp %>% pivot_longer(.,!state,names_to = "year",values_to = "gdp")
gdp_l <- gdp_l %>% mutate(.,st_yr=paste0(state,year))

FFR <- read_csv("FFR.csv")
FFR$year <- as.factor(FFR$year)

B_tax <- policy_tax[,3:4]
colnames(B_tax) <- c("st_yr","B_pol_tax")

controls <- left_join(num_hs_l,pop_l[,3:4],by=c("st_yr"))
controls <- left_join(controls,gdp_l[,3:4],by=c("st_yr"))
controls <- left_join(controls,FFR[,c(1,4)],by=c("year"))
controls <- left_join(controls,B_tax,by=c("st_yr"))

scale6<- function(x){x=x/1000000}
controls <- controls%>%mutate_at(c("pop","num_house","gdp"),scale6)

controls[is.na(controls)]<-0
```

## merge files
```{r}
# merge and format files
data_panel <- HPI_yr_l[which(HPI_yr_l$abbr!="DC"),]
data_panel <- left_join(data_panel,HPI_pol_lag_l[,c(3,4)],by=c("st_yr"))
data_panel <- left_join(data_panel,HPI_nbrnpol_l[,c(3,4)],by=c("st_yr"))
data_panel <- left_join(data_panel,tax_pol_lag_l[,c(3,4)],by=c("st_yr"))
data_panel <- left_join(data_panel,bank_pol_lag_l[,c(3,4)],by=c("st_yr"))
data_panel <- left_join(data_panel,const_pol_lag_l[,c(3,4)],by=c("st_yr"))
data_panel <- left_join(data_panel,controls[,c(3:8)],by=c("st_yr"))
data_panel <- left_join(data_panel,bank_nbr_lag_l[,c(3,4)],by=c("st_yr"))
data_panel <- left_join(data_panel,tax_nbr_lag_l[,c(3,4)],by=c("st_yr"))
data_panel <- left_join(data_panel,const_nbr_lag_l[,c(3,4)],by=c("st_yr"))
data_panel <- left_join(data_panel,emp_pol_lag_l[,c(3,4)],by=c("st_yr"))
data_panel <- left_join(data_panel,emp_nbr_l[,c(3,4)],by=c("st_yr"))

pdata_panel <- pdata.frame(data_panel)

# creating lags with plm::lag()
z <- pdata_panel$rate
z <- lag(z)
pdata_panel$rate_lag1<-z

z <- pdata_panel$HPI_pol_lag
z <- lag(z)
pdata_panel$HPI_pol_lag1<-z

z <- pdata_panel$HPI_nbrnpol_lag
z <- lag(z)
pdata_panel$HPI_nbrnpol_lag1<-z

#save(pdata_panel,file="pdatapanel.RData")
```

# the regression models

## panel, not time autoregressive, within estimator  
```{r}
plm_dyn_1 <- plm(rate~HPI_nbrnpol_lag1+num_house+pop+gdp+lag(emp,1),data=pdata_panel,index=c("abbr","year"),model="within")
plm_dyn_3 <- plm(rate~HPI_pol_lag1+num_house+pop+gdp+lag(emp,1),data=pdata_panel,index=c("abbr","year"),model="within")
plm_dyn_5 <- plm(rate~HPI_nbrnpol_lag1+HPI_pol_lag1+num_house+pop+gdp+lag(emp,1),data=pdata_panel,index=c("abbr","year"),model="within")

stargazer(plm_dyn_1,plm_dyn_3,plm_dyn_5,type="text",title="dynamic panel, within estimator",dep.var.caption="yearly HPI growth rate",no.space=TRUE,omit.stat=c("LL","ser","F"),column.sep.width="3pt")

```


## for r squared. panel, with time autoregressive rate_t-1, within estimator 
```{r}
plm_dyn2_1 <- plm(rate~rate_lag1+HPI_nbrnpol_lag1+num_house+pop+gdp+lag(emp,1),data=pdata_panel,index=c("abbr","year"),model="within")
plm_dyn2_3 <- plm(rate~rate_lag1+HPI_pol_lag1+num_house+pop+gdp+lag(emp,1),data=pdata_panel,index=c("abbr","year"),model="within")
plm_dyn2_4 <- plm(rate~rate_lag1+num_house+pop+gdp+lag(emp,1),data=pdata_panel,index=c("abbr","year"),model="within")
plm_dyn2_5 <- plm(rate~rate_lag1+HPI_nbrnpol_lag1+HPI_pol_lag1+num_house+pop+gdp+lag(emp,1),data=pdata_panel,index=c("abbr","year"),model="within")

stargazer(plm_dyn2_4,plm_dyn2_1,plm_dyn2_3,plm_dyn2_5,type="text",title="dynamic panel regression, all key regressors time lagged",dep.var.caption="yearly HPI growth rate",no.space=TRUE,omit.stat=c("LL","ser","F"),column.sep.width="3pt")
```



## dynamic panel, sysGMM estimator
```{r}
# GMM uses all time laggs to instrument for the endogenous regressor
# formula specified in two parts, regression formula in part one, GMM instruments for the endogenous regressor in part two, optional part three puts the "normal" instruments
# transformation="d" for difference GMM, ld" for sysGMM estimator
# model= "individual", or "twoways"

plm_gmm_1 <- pgmm(rate~lag(rate,1)+HPI_nbrnpol_lag1+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2)+lag(HPI_nbrnpol_lag,2),data=pdata_panel,effect="individual",index=c("abbr","year"),model="onestep",transformation="ld")

plm_gmm_3 <- pgmm(rate~lag(rate,1)+HPI_pol_lag1+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2)+lag(HPI_pol_lag1,1),data=pdata_panel,effect="individual",index=c("abbr","year"),model="onestep",transformation="ld")

plm_gmm_4 <- pgmm(rate~lag(rate,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2),data=pdata_panel,effect="individual",index=c("abbr","year"),model="onestep",transformation="ld")

plm_gmm_5 <- pgmm(rate~lag(rate,1)+HPI_nbrnpol_lag1+HPI_pol_lag1+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2)+lag(HPI_nbrnpol_lag,2)+lag(HPI_pol_lag1,1:30),data=pdata_panel,effect="individual",index=c("abbr","year"),model="onestep",transformation="ld")

stargazer(plm_gmm_4,plm_gmm_1,plm_gmm_3,plm_gmm_5,type="text",title="dynamic panel, sysGMM estimator",dep.var.caption="yearly HPI growth rate",no.space=TRUE,omit.stat=c("LL","ser","F"),column.sep.width="3pt")

```


```{r}
# get robust standard errors with package clusterSEs
cluster.bs.plm(mod=plm_gmm_1,dat=pdata_panel,cluster="group",ci.level=0.95,boot.reps=1000,cluster.se=TRUE,report=TRUE,prog.bar=TRUE)
```


## dynamic panelwith IV, sysGMM estimator 
instrument for HPI_nbrnpol_lag1 with bank_nbr_lag, emp_nbr_lag
instrument for HPI_pol_lag1 with tax_pol_lag, emp_pol_lag

```{r}
# tax+ const
plm_gmm_iv1 <- pgmm(rate~lag(rate,1)+HPI_nbrnpol_lag1+HPI_pol_lag1+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2:30)|tax_nbr_lag+const_pol_lag+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata_panel,effect="individual",index=c("abbr","year"),model="onestep",transformation="ld")

# bank + emp
plm_gmm_iv2 <- pgmm(rate~lag(rate,1)+HPI_nbrnpol_lag1+HPI_pol_lag1+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2:30)|bank_nbr_lag+lag(emp_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata_panel,effect="individual",index=c("abbr","year"),model="onestep",transformation="ld")

# emp + emp
plm_gmm_iv3 <- pgmm(rate~lag(rate,1)+HPI_nbrnpol_lag1+HPI_pol_lag1+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2:30)|lag(emp_nbr_lag,1)+lag(emp_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata_panel,effect="individual",index=c("abbr","year"),model="onestep",transformation="ld")

# lag + emp
plm_gmm_iv4 <- pgmm(rate~lag(rate,1)+HPI_nbrnpol_lag1+HPI_pol_lag1+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2:30)+lag(HPI_nbrnpol_lag,2:30)|lag(emp_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata_panel,effect="individual",index=c("abbr","year"),model="onestep",transformation="ld")

stargazer(plm_gmm_iv1,plm_gmm_iv2,plm_gmm_iv3,plm_gmm_iv4,type="text",title="dynamic panel wth IV, sysGMM estimator",dep.var.caption="yearly HPI growth rate",no.space=TRUE,omit.stat=c("LL","ser","F"),column.sep.width="3pt")

```


## for r squared

```{r}
# tax+ const
plm_iv1a <- plm(rate~lag(rate,1)+HPI_nbrnpol_lag1+HPI_pol_lag1+tax_nbr_lag+const_pol_lag+log(num_house)+log(pop)+log(gdp)+lag(emp,1), data=pdata_panel,index=c("abbr","year"),model="within")

# bank + emp
plm_iv2a <- plm(rate~lag(rate,1)+HPI_nbrnpol_lag1+HPI_pol_lag1+bank_nbr_lag+lag(emp_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata_panel,index=c("abbr","year"),model="within")

# emp + emp
plm_iv3a <- plm(rate~lag(rate,1)+HPI_nbrnpol_lag1+HPI_pol_lag1+lag(emp_nbr_lag,1)+lag(emp_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata_panel,index=c("abbr","year"),model="within")

# lag + emp
plm_iv4a <- plm(rate~lag(rate,1)+HPI_nbrnpol_lag1+HPI_pol_lag1+lag(HPI_nbrnpol_lag1)+lag(emp_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata_panel,index=c("abbr","year"),model="within")

stargazer(plm_iv1a,plm_iv2a,plm_iv3a,plm_iv4a,type="text",title="dynamic panel wth IV, sysGMM estimator",dep.var.caption="yearly HPI growth rate",no.space=TRUE,omit.stat=c("LL","ser","F"),column.sep.width="3pt")

```



## subperiods 2000-2007, 2012-2019

```{r}

pdata1 <- data_panel[which(as.numeric(data_panel$year)<2008),]
pdata2 <- data_panel[which(as.numeric(data_panel$year)>2010 & as.numeric(data_panel$year)<2019),]
```



```{r}
# dynamic panel, sysGMM estimator

# 2000-2008 
plm_gmm_iv1b <- pgmm(rate~lag(rate,1)+lag(HPI_nbrnpol_lag,1)+lag(HPI_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2:30)|lag(emp_nbr_lag,1)+lag(emp_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata1,effect="individual",index=c("abbr"),model="onestep",transformation="d")

# 2012-2019
plm_gmm_iv2b <- pgmm(rate~lag(rate,1)+lag(HPI_nbrnpol_lag,1)+lag(HPI_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2:30)|lag(emp_nbr_lag,1)+lag(emp_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata2,effect="individual",index=c("abbr"),model="onestep",transformation="d")

# 2000-2008 
plm_gmm_iv3b <- pgmm(rate~lag(rate,1)+lag(HPI_nbrnpol_lag,1)+lag(HPI_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2:30)|tax_nbr_lag+const_pol_lag+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata1,effect="individual",index=c("abbr"),model="onestep",transformation="d")

# 2012-2019
plm_gmm_iv4b <- pgmm(rate~lag(rate,1)+lag(HPI_nbrnpol_lag,1)+lag(HPI_pol_lag,1)+log(num_house)+log(pop)+log(gdp)+lag(emp,1) | lag(rate,2:30)|tax_nbr_lag+const_pol_lag+log(num_house)+log(pop)+log(gdp)+lag(emp,1),data=pdata2,effect="individual",index=c("abbr"),model="onestep",transformation="d")

stargazer(plm_gmm_iv1b,plm_gmm_iv2b,plm_gmm_iv3b,plm_gmm_iv4b,type="text",title="dynamic panel, sysGMM estimator",dep.var.caption="yearly HPI growth rate",no.space=TRUE,omit.stat=c("LL","ser","F"),column.sep.width="3pt")

#cluster.bs.plm(mod=plm_gmm_iv3b,dat=pdata2,cluster="group",ci.level=0.95,boot.reps=1000,cluster.se=TRUE,report=TRUE,prog.bar=TRUE)
```

## examin collinearity
```{r}
ggplot(pdata_panel,aes(x=HPI_pol_lag,y=HPI_nbrnpol_lag,color=year))+
  geom_point()
ggplot(pdata_panel,aes(y=HPI_pol_lag,x=bank_pol_lag,color=year))+
  geom_point()

cor(pdata_panel$HPI_pol_lag,pdata_panel$HPI_nbrnpol_lag) # 0.805
cor(pdata_panel$rate,pdata_panel$HPI_nbrnpol_lag) # 0.83
cor(pdata_panel$rate,pdata_panel$HPI_pol_lag) # 0.72

cor(pdata_panel$tax_pol_lag,pdata_panel$bank_pol_lag) # 0.53
cor(pdata_panel$HPI_pol_lag,pdata_panel$const_pol_lag) # 0.08

```

```{r}
ggplot(pdata_panel,aes(y=rate,x=HPI_nbrnpol_lag,color=year))+
  geom_point()

ggplot(pdata_panel,aes(y=rate,x=HPI_pol_lag,color=year))+
  geom_point()
```








